A Data-Driven Framework for Neural Field Modeling

### 2011-01-18 ltx/first_order_framework/response_to_reviews.tex

	\emph{A more sophisticated model for the error will provide a more realistic description of the cortical dynamics. However, this will increase the number of parameters making the estimation problem more difficult. Furthermore, the authors believe that is a benefit staying as close as possible to the classic WC equations, so future results can be compared to theoretical studies. A preferable alternative to an AR error term is to allow for time-varying parameters to account for changes in excitability levels. In future work, we also hope estimate the disturbance properties (spatial covariance and temporal variance). To address this point in this paper, we have added an extra paragraph to `Neural Field Model' section of the Methods, where we discuss the model mismatch and some basic modeling assumptions, such as stationarity of the parameters. We have also added to the Estimation section, stating that we are assuming the parameters are stationary of the estimation period. In addition, we have added to the discussion (Extensions section) with possible ways to relax these assumptions.}

\item Although we all know the 1970s version of Wilson and Cowan's field equations, there was an intriguing stochasticization of these equations by Benayoun et al PloS Comp Biol 2010. I and others will value the authors' thoughts of how a more palatable stochasticization of these field equations might influence the future development of the data-assimilation framework presented in this paper.

\emph{We have attempted to answer this assuming the reviewer is after a sensible method for choosing the disturbance parameters.}

\emph{The results of Benayoun et al. provide evidence that the variance in a stochastic WC model will depend on the number of neurons that are modeled (in a stochastic network), where a larger network behaves more like the deterministic WC equations. This work encourages the use of the WC model with an additive disturbance for data-assimilation, but suggests that the disturbance properties should also be estimated. The authors believe that estimating the noise properties is possible within an EM framework, and we are planning to demonstrate how to achieve this in our future work. We have added some comments to the Extension to the Framework section in this regard.}

### 2011-01-18 ltx/first_order_framework/response_to_reviews.tex

\item I understand that the Mexican Hat bases are in this paper often a composite of $\theta_0$, $\theta_1$, and $\theta_2$, representing short range excitation, short range inhibition, and long range excitation. But many will be confused by this subtlety, and at times in the text the Mexican Hat basis is described as if it were a single function. I would show how the theta's sum to create the general spatial basis shape of the hat in figure 1, and go through the paper in order to clarify this issue.

\emph{This point is clarified by adding the connectivity kernel components superimposed by the resulting Mexican Hat basis in Figure 1. of the manuscript. It has also been clarified throughout the paper.}
 % 
 % \begin{figure*}[!ht]
 % \begin{center}
 % \includegraphics{./Graph/pdf/fig1.pdf} 
 % \end{center}
 % \caption{}
 % \label{fig:Figure1}
 % \end{figure*}

\item A few comments on how to adapt the smoother for real time implementation would be helpful. I presume that one would perform the smoothing during a training phase. But one would also wish to adapt in real time as observation or control scenarios progressed. A few words of perspective would be invaluable to help others translate the author's wisdom on this method to other applications.

\emph{A few comments have been added to the last paragraph of the `Extensions to the Framework' section of the discussion.}

### 2011-01-18 ltx/first_order_framework/LaTeX_Manuscript.tex
got rid of colour in the filter
minor changes

The additive form of the unscented Rauch-Tung-Striebel smoother (URTSS) \citep{Sarkka2010} was used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF) \citep{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{align}\label{eq:sigmapoints1}
	\mathcal X_{0}&=\mathbf{\bar x} \\
	\mathcal X_{i}&= \mathbf{\bar x}+\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, n_x \\
	\mathcal X_{i}&=\mathbf{\bar x}-\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_{i- n_x}, \quad i= n_x+1, \dots, 2n_x 
\end{align}
where $\mathbf{\bar x}$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for filtering or smoothing, $\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$, and $n_x$ is the dimension of the state space. The total number of sigma points is $2n_x+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( n_x+\kappa) - n_x, 
\end{equation}
The positive constant $\alpha$, that determines the spread of the sigma points around $\mathbf{\bar x}$, was set to $10^{-3}$. It is typically set arbitrary small to minimise higher-order effects \citep{Haykin2001}. The other scaling parameter, $\kappa$, was set to $3-n_x$ in accordance to \citet{Julier2002a}. \dean{Parham to add something about the sensitivity of the parameters.}

@@

The estimation framework that we have presented is inherently a batch algorithm designed to operate after data has been collected. A further extension of the framework is to allow for the algorithm to learn online. A filter that is designed to be implemented in an online setting can be used in place of the smoother along with an online parameter estimator to update the parameter after a new observation arrived. A real-time implementation can be achieved using the UKF (instead of the URTSS) and an online recursive LS estimator \cite{Ljung1999}.

>>>>

The additive form of the unscented Rauch-Tung-Striebel smoother (URTSS) \citep{Sarkka2010} was used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF) \citep{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{align}\label{eq:sigmapoints1}
	\mathcal X_{0}&=\mathbf{\bar x} \\
	\mathcal X_{i}&= \mathbf{\bar x}+\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, n_x \\
	\mathcal X_{i}&=\mathbf{\bar x}-\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_{i- n_x}, \quad i= n_x+1, \dots, 2n_x 
\end{align}
where $\mathbf{\bar x}$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for filtering or smoothing, $\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$, and $n_x$ is the dimension of the state space. The total number of sigma points is $2n_x+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( n_x+\kappa) - n_x, 
\end{equation}
The positive constant $\alpha$, that determines the spread of the sigma points around $\mathbf{\bar x}$, was set to $10^{-3}$. It is typically set arbitrary small to minimise higher-order effects \citep{Haykin2001}. The other scaling parameter, $\kappa$, was set to $3-n_x$ in accordance to \citet{Julier2002a}.

@@

The estimation framework that we have presented is inherently a batch algorithm designed to operate after data has been collected. A further extension of the framework is to allow for the algorithm to learn online. A filter that is designed to be implemented in an online setting can be used in place of the smoother along with an online parameter estimator to update the parameters after a new observation arrived. A real-time implementation can be achieved using the UKF (instead of the URTSS) and an online recursive LS estimator \cite{Ljung1999}.
